---
title: "Using fmisc Custom Linting Rules"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using fmisc Custom Linting Rules}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

The `fmisc` package provides custom linting rules for the **flir** package. This vignette demonstrates how to use these custom rules in your R projects.

# Using flir Custom Rules

## Setup

First, set up flir in your project:

```{r eval=FALSE}
library(flir)
setup_flir()
```

This creates a `flir/` directory with a `config.yml` file.

## Loading fmisc Rules

Edit `flir/config.yml` to include fmisc rules:

```yaml
from-package:
  - fmisc
```

You can also manually get the rules path:

```{r eval=FALSE}
library(fmisc)
get_flir_rules()
```

## Running flir

Check your code for issues:

```{r eval=FALSE}
# Lint all R files in current directory
flir::lint_dir()

# Lint specific files
flir::lint_file("my_script.R")
```

Apply automatic fixes:

```{r eval=FALSE}
# Fix all R files in current directory
flir::fix_dir()

# Fix specific files
flir::fix_file("my_script.R")
```

## Available flir Rules

### deprecated-sample-n

Updates dplyr's deprecated `sample_n()`:

**Before:**
```r
sampled <- sample_n(mtcars, 10)
```

**After:**
```r
sampled <- slice_sample(mtcars, n = 10)
```

### deprecated-sample-frac

Updates dplyr's deprecated `sample_frac()`:

**Before:**
```r
sampled <- sample_frac(mtcars, 0.1)
```

**After:**
```r
sampled <- slice_sample(mtcars, prop = 0.1)
```

# Creating Custom flir Rules

Create a YAML file (e.g., `my-rule.yml`):

```yaml
id: my-custom-rule
language: r
severity: warning
rule:
  pattern: old_function($ARG1, $ARG2)
message: "Use new_function() instead of old_function()"
fix: new_function(~~ARG1, ~~ARG2)
```

### Rule Structure

- `id`: Unique identifier for the rule
- `language`: Programming language (always "r" for R)
- `severity`: One of "warning", "error", or "info"
- `rule`: Pattern matching specification
  - `pattern`: ast-grep pattern to match
  - `any`: List of alternative patterns
  - `inside`: Context where pattern should match
- `message`: Description shown to users
- `fix`: (Optional) Automatic replacement

### Using Metavariables

Metavariables capture code patterns:

- In patterns: Use `$VAR` syntax
- In messages/fixes: Use `~~VAR` syntax

Example with multiple patterns:

```yaml
id: modernize-apply
language: r
severity: warning
rule:
  any:
    - pattern: sapply($DATA, $FUN)
    - pattern: sapply($DATA, $FUN, simplify = $SIMP)
message: "Consider using vapply() for type-stable results"
```

For more information:

- [flir documentation](https://flir.etiennebacher.com/)
- [fmisc GitHub repository](https://github.com/fabiandistler/fmisc)
