---
title: "Function Decorators and the R Ecosystem"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Function Decorators and the R Ecosystem}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

`fmisc` ships a small toolkit of composable function operators
("decorators"): `with_retry()`, `with_timing()`, `with_logging()`,
`with_cache()`, and `with_rate_limit()`. They cover the common
cross-cutting concerns without pulling in heavy dependencies, compose
freely with the native pipe, and can be inspected (`print()`),
unwrapped (`undecorate()`), and tested in isolation.

They intentionally stay minimal. This vignette summarises when to keep
using them and when to reach for a dedicated package instead.

# Decorators vs. the ecosystem

## Retries: `with_retry()` vs `purrr::insistently()`

Use `with_retry()` when you need exponential backoff with jitter,
selective retry via `.on_error`, and a retry callback — with no
dependencies beyond `cli`.

Reach for `purrr::insistently()` (or `purrr::quietly()` /
`purrr::safely()`) when you are already inside a `purrr` pipeline and
want retry semantics that integrate with its adverb style. Note that
`insistently()` retries on *any* error unless combined with
`rate_backoff()`, and does not offer exponential backoff with jitter by
itself.

## Caching: `with_cache()` vs `memoise` + `cachem`

There is no contest here — because `with_cache()` *is* a facade over
`memoise::memoise()` backed by `cachem::cache_mem()`:

```{r eval=FALSE}
fast <- with_cache(slow, .max_size = 1024, .ttl = 3600)
```

`.max_size` maps to `cachem`'s `max_n` (LRU eviction), `.ttl` to
`max_age`. If you need more than an in-memory LRU/TTL cache — disk
persistence (`cache_disk`), shared caches, or cache pruning — use
`memoise` directly; every `cachem` backend is one argument away:

```{r eval=FALSE}
memo <- memoise::memoise(slow, cache = cachem::cache_disk("~/.cache"))
```

## Rate limiting: `with_rate_limit()` vs `ratelimitr`

`with_rate_limit()` enforces at most `n` calls per rolling `period`
seconds, either sleeping (`.wait = TRUE`) or raising
`"fmisc_rate_limit_exceeded"`.

For multi-function rate limiting across a group of calls, token-bucket
semantics, or self-resetting rate-limited functions with finer control,
use `ratelimitr::limit_rate()`.

## HTTP-specific throttling

**Burst-tolerant rate limiting is deliberately not implemented in
fmisc.** The strict rolling window is intentional: it is predictable,
easy to reason about, and matches how most API quotas are documented.
If you need burst tolerance for HTTP traffic (allowing short spikes up
to a bucket capacity while staying under a long-run average), do not
reach for this decorator — use `httr2::req_throttle()`, which implements
a proper token-bucket scheme and integrates with httr2's request
pipeline:

```{r eval=FALSE}
request("https://api.example.com") |>
  req_throttle(rate = 10 / 1) |>
  req_perform()
```

# Inspecting and unwrapping

Decorators record what was applied, so decorated functions explain
themselves:

```{r}
f <- function(x) x + 1
g <- f |>
  fmisc::with_retry(max_tries = 3L, backoff = 0) |>
  fmisc::with_timing(.report = "attribute")
print(g)
```

Peel layers off with `undecorate()`: one layer by default, or all of
them with `depth = Inf`.

```{r}
identical(fmisc::undecorate(g, depth = Inf), f)
```
